Title

### 2013-03-25 inc/design_considerations/considerations.tex
rather large number of tweaks to the considerations section

In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effect that column elements have upon it.
However, microscopy is an experimental science, and designing a UEM is no less hands-on.

@@

These considerations drive the remainder of this thesis.
In this chapter, I will first convince the reader of these considerations.
As a direct consequence of these considerations, it is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
In the following chapters I will show how we are addressing these considerations for the prototype system at UIC.

@@

In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode.
Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially non-directional.
A pinhole is used to eliminate most of the electrons with the exception of those which will travel down the column, reducing the beam current but producing a high-quality beam. %TODO ref(s)

In fact, for conventional electron microscopes without time resolution, the Rose criterion may be met for even very low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
Further, high beam current may be damaging to the sample over time, so often a lower beam current is preferred.

By definition, for single-shot imaging with a DTEM, all of these electrons must be contained in a single pulse.
Although sample damage concerns may be neglected, destructive measurements will result in useful images, the use of a pinhole is likely to reduce the pulse charge below the useful level.
This means that the designer of a DTEM must endevor to keep all of the created electrons.

It should be noted that for diffraction studies with a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.

@@

This steady-state law for the maximum current density that can be achieved is not valid in the regime of interest for UEM (and UED) where the use of picosecond and femtosecond photocathode drive laser pulses implies that $t_{gun}$ is usually much greater than the electron pulse duration $\tau_e$.
In fact, though there is no clear definition in the literature or the community, we propose that this is useful delimination between term UEM and the more general term DTEM; reserving the term UEM for when the pulse can be completely contained inside the accelerator, i.e. $t_{gun} > \tau_e$.
In this short pulse case, an equivalent diode approximation \cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
\begin{equation}
  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL}
\end{equation}
where $\chi = \tau_e / t_{gun}$.
Thus, for the usual case in UEM, where $\chi \ll 1$, the critical current density is simply given by
\begin{equation}
  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e}
\end{equation}
where the DC gun field $E_{DC} = V_{DC}/d$.
This relation may be rewritten to yield a critical number of photoemitted electrons/pulse before virtual cathode formation:
\begin{equation} \label{eq:n_crit}
  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d}
\end{equation}
In obtaining the above equation, we have used an irradiated photocathode area of $\pi w^2$ associated with an incident Gaussian laser pulse of the form of \ref{eq:laser_form}.
The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.

@@

This finally leads to the first consideration for UEM design mentioned in the introduction: for the clean pulsed emission of sufficient to create a useful image, the generated beam should have a HW1/eM width 
\begin{equation} \label{eq:minimum_w}
  w \gtrsim \sqrt{\frac{N e d}{\pi \varepsilon_0 V_{DC}}} \,\text{.}
\end{equation}
For our example case of a 1k $\times$ 1k CCD, $N \ge 10^8$ electrons/pulse, we require $w \gtrsim $ 0.56 mm.

@@

This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred.
Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
Though deriving an absolute limit on the allowable transverse emittance for imaging in a UEM would depend on many column-specific factors, since diffraction is expected to impose less stringent requirements on the beam than imaging, we may consider it when searching for an upper bound on transverse emittance.

@@

>>>>

In Chapters \ref{chap:previous_models} through \ref{chap:model_results}, I have explored the mathematical side of designing an ultrafast electron microscope; modeling the behavior of a pulse and the effects that column elements have upon it.
However, microscopy is an experimental science, and designing a UEM is no less hands-on.

@@

These considerations drive the remainder of this thesis.
In this chapter, I will first convince the reader of these considerations.
As a direct consequence of these considerations, it is our belief that a successful UEM project cannot simply stem from small modifications to the platform of a conventional electron microscope.
In the following chapters I will show how we are addressing these considerations for the prototype system at UIC, both in column design (Chapter \ref{chap:prototype}) and photocathode engineering (Chapter \ref{chap:photocathode}).

@@

In a conventional electron microscope the electrons are generated by one of several processes, often by heating the cathode or applying a large field to the cathode, or combinations thereof.
Although these processes produce a very large number of electrons, far more than required by the Rose criterion, they are essentially non-directional.
A pinhole is used to eliminate most of the electrons with the exception of those which will travel down the column, reducing the beam current but producing a high-quality beam. %TODO ref(s)

In fact, for conventional electron microscopes without time resolution, the Rose criterion may be met for even very low beam current by simply allowing the image to acquire for a long enough time to collect sufficent electrons.
Further, high beam current may be damaging to the sample over time, so a lower beam current is often preferred.

By definition, for single-shot imaging with a DTEM, all of these electrons must be contained in a single pulse.
Although sample damage concerns may be neglected, destructive measurements will result in useful images, the use of a pinhole is likely to reduce the pulse charge below the useful level.
This means that the designer of a DTEM must endeavor to retain and utilize all of the generated electrons.

It should be noted that for diffraction studies performed in a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.

@@

This steady-state law for the maximum current density that can be achieved is not valid in the regime of interest for UEM (and UED) where the use of picosecond and femtosecond photocathode drive laser pulses implies that $t_{gun}$ is usually much greater than the electron pulse duration $\tau_e$.
Though there is no clear delineation in the literature or the community between moniker of Ultrafast Electron Microscopy (UEM) and the more general term Dynamic Transmission Electron Microcopy (DTEM), we propose that this is useful delimination, reserving the term UEM for those systems where the pulse can be completely contained inside the accelerator, i.e. $t_{gun} > \tau_e$.
Pulses of this type are subject to a different ``short pulse analog'' of the Child-Langmuir limit.

In this short pulse case, an equivalent diode approximation \cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
\begin{equation}
  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL}
\end{equation}
where $\chi = \tau_e / t_{gun}$.
When $\chi \ll 1$, the usual case in UEM, the critical current density is simply given by
\begin{equation}
  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e}
\end{equation}
where the DC gun field $E_{DC} = V_{DC}/d$.
This relation may be rewritten to yield a critical number of photoemitted electrons/pulse before virtual cathode formation:
\begin{equation} \label{eq:n_crit}
  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d}
\end{equation}
In obtaining the above equation, we have used an irradiated photocathode area of $\pi w^2$ associated with an incident Gaussian laser pulse of the form of \ref{eq:laser_form}.
The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.

@@

This finally leads to the first consideration for UEM design mentioned in the introduction: for the clean pulsed emission of sufficient to create a useful image, the generated beam should have a HW1/eM width 
\begin{equation} \label{eq:minimum_w}
  w \gtrsim \sqrt{\frac{N e d}{\pi \varepsilon_0 V_{DC}}} \,\text{.}
\end{equation}
For our example case above, to cleanly extract $N \ge 10^8$ electrons/pulse we require $w \gtrsim 0.56 \text{ mm}$.
I will discuss the effect that accomodating a large pulse size has on designing an Ultrafast Electron Microscope in Chapter \ref{chap:prototype}.

@@

This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
Though deriving an absolute limit on the allowable transverse emittance for imaging in a UEM would depend on many column-specific factors, since diffraction is expected to impose less stringent requirements on the beam than imaging, we may consider it when searching for an upper bound on transverse emittance.

@@

To have $\Delta X = $1 nm spatial resolution, and be able to measure a 10\% change ($\Delta a / a = 0.1$) in a Bragg spacing $a \sim$ 1-2 \AA{} in the 1$^{st}$ diffraction order, the example DTEM from Section \ref{sec:childs_law}, equipped with a 1k x 1k CCD will require $\varepsilon_{\smallT} \lesssim$ 2 nm.
For comparison, using the simplistic initial conditions in \ref{eq:summary}, exciting carriers from a HW1/eM $w = 0.56 \text{ mm}$ Gaussian spot-size on a Tantalum metal photocathode (work function $\Phi =$ 4.25 eV) using our laser (photon energy $\hbar \omega =$ 4.75 eV, further described in Section \ref{sec:laser}), for an excess photoemission energy $\Delta E = \hbar \omega - \Phi =$ 0.5 eV, the normalized transverse emittance is 226 nm.

One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, reducing the excess photoemission energy $\Delta E$.
Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.
In Chapter \ref{chap:photocathode}, I will explore several more exotic photoemission processes in an attempt to reduce $\varepsilon_{\smallT}$ without this penalty.
